---
title: 'analysis of Pterostichus madidus morphology on 2004-2005 ground beetle data'
author: ""
date:
output:
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
library(tidyverse)
library(sf)
library(lubridate)
library(brms)
library(cmdstanr)
library(tidybayes)

library(here)

options(mc.cores = 4) 
```

# Introduction

<!--SOME CONTEXT TO ADD -->
As far as I can tell, nobody really used *Pterostichus madidus* polymorphism as an evolutionary ecology model in cities (not a lot of people used it as an evolutionary model *at all* actually, even outside cities).
There's a few urban data in that 2002 thesis, but a bit messy; there is also an actual *P. madidus* urban evolutionary study (carabids more generally actually; the refs are in the urban evolution book), but it was based on genetic diversity, not phenotypic polymorphism, so we're safe.

# Data loading and preparation

First we load all the data (note that *P. madidus* data come in two separate files, one for the red-legged morph, one for the black-legged).

```{r data}
raw_counts <-read_csv(here("data","raw_data","pterostichus_madidus_2004_2005.csv"))
dates <- read_csv(here("data","raw_data","sampling_dates_2004_2005.csv"))

urban_info <- read_csv(here("data","processed_data","urban_info_IMD.csv"))


raw_sites <-read_sf(here("data","GIS_layers","rennes2004_sites_centroids.gpkg"),layer="rennes2004_sites_centroids")
ucdb_centroid <- read_sf(here("data","GIS_layers","rennes_ucdb_2015_centroid.gpkg"),layer="rennes_ucdb_2015_centroid")
ucdb_extent <- read_sf(here("data","GIS_layers","rennes_ucdb_2015_extent.gpkg"),layer="rennes_ucdb_2015_extent")
```

<!--describe data content-->

Now we reshape the data so they are in the long format, we remove the `RB12` site which was more or less abandoned because boars (see Solène's paper), and we add and reorder some variables so that things are easier down the line

```{r}
summed_counts <- raw_counts |> 
  rowwise() |> 
  mutate(N = sum(c_across(P1:P12))) |> 
  ungroup() |> 
  select(CODE,CAMP,SITE,N) |> 
  pivot_wider(names_from=CODE,values_from = N) |> 
  rename(N_blackleg="PTEMN",N_redleg="PTEMR") |> 
  mutate(N_total=N_blackleg+N_redleg)
```


```{r}
sites <- raw_sites |> 
  rename(SITE="site") |> 
  mutate(dist_urban_centroid_m=as.numeric(st_distance(raw_sites,ucdb_centroid)[,1])) |> 
  left_join(urban_info) |> 
  mutate(scaled_dist = scale(dist_urban_centroid_m)[,1], ##scaled to facilitate model fitting
         scaled_IMD_100 = scale(meanIMD_100m)[,1],
         scaled_IMD_300 = scale(meanIMD_300m)[,1],
         scaled_IMD_600 = scale(meanIMD_600m)[,1],
         scaled_IMD_900 = scale(meanIMD_900m)[,1],
         scaled_IMD_1200 = scale(meanIMD_1200m)[,1],
         scaled_IMD_1500 = scale(meanIMD_1500m)[,1],
         scaled_IMD_1800 = scale(meanIMD_1800m)[,1])
```

```{r}
data <- summed_counts |> 
  filter(N_total>0)  |>  #we remove all dates where no madidus at all were found
  left_join(sites)  |> 
  left_join(dates)  |>  
  mutate(SPECIES="Pterostichus madidus",
         doy_start=yday(dmy(DATE_start)),
         doy_end=yday(dmy(DATE_end)),
         YEAR=factor(YEAR))  |>  
  filter(SITE!="RB12")  ## that site should be removed because trampled by boars (see Solène paper)
```

We note a few numbers useful for the Material and Methods:
```{r}
sum(summed_counts$N_total) # total beetles in the source datasets (so including RB12)

sum(data$N_total) # total beetles in the final dataset (so excluding RB12)
```

We can do a rough visualisation
```{r}
data  |>  
  ggplot()+
  geom_point(aes(dist_urban_centroid_m,N_blackleg/N_total,
                 size=N_total,col=YEAR))


data  |>  
  ggplot()+
  geom_point(aes(meanIMD_100m,N_blackleg/N_total,
                 size=N_total,col=YEAR))+facet_wrap(~CAMP)
```

mm, looks like there is clearly an effect of urbanisation
# Models

Our model assumes an effect of urbanisation (fixed effects), as well as other site-level variation or the possibility that the average frequency of morphs and the effect of urbanisation may change through time (due to seasonality) through random effects.

priors are relatively standard, especially since we scale our urbanisation variables

```{r}
priorB=c(
          set_prior("normal(0,1)",class="b"),
          set_prior("normal(0,1.5)",class="Intercept"),
          set_prior("normal(0,1)",class="sd")
        )
```

```{r}
if (file.exists(here("R_output", "model_distanceB.RDS")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the RDS object)
  {
    mod_distB <- readRDS(here("R_output", "model_distanceB.RDS"))
  } else {
    
mod_distB <- brm(
  bf(N_blackleg|trials(N_total)~scaled_dist+
                                      (1|SITE)+ (scaled_dist | CAMP)),
        family=binomial,
        data=data,
        backend="cmdstanr",
        seed=42,prior=priorB
        )

  saveRDS(mod_distB, file = here("R_output", "model_distanceB.RDS"))
  }
```


```{r}
pp_check(mod_distB,"stat_2d") ##possibly some slight overdispersion

loo_distB<-loo(mod_distB) # warning about some problematic observations, so we could use
# loo_distB2<-loo(mod_distB,moment_match = TRUE) to reduce them
#but does not reduce them fully, so might be best to use K-fold CV for everything since we will need to refit the models a certain amount of time

if (file.exists(here("R_output", "KfoldCV_distB.RDS"))){
  KCV_distB <- readRDS(here("R_output","KfoldCV_distB.RDS"))
  }else{
    set.seed(42)
    KCV_distB<-kfold(mod_distB,K=10)
    saveRDS(KCV_distB,file=here("R_output","KfoldCV_distB.RDS"))
    }
```

then we do the same for a beta-binomial model

```{r}
priorBB=c(
          set_prior("normal(0,1)",class="b"),
          set_prior("normal(0,1.5)",class="Intercept"),
          set_prior("normal(0,1)",class="sd"),
          set_prior("normal(0,1)",nlpar="invphi",lb=0)
        )
```


```{r}

if (file.exists(here("R_output", "model_distanceBetaB.RDS")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the RDS object)
  {
    mod_distBB <- readRDS(here("R_output", "model_distanceBetaB.RDS"))
  } else {
mod_distBB <- brm(
  bf(N_blackleg|trials(N_total)~scaled_dist+
                                      (1|SITE)+ (scaled_dist | CAMP),
     nlf(phi~1/invphi),
     invphi~1),
        family=beta_binomial(link_phi="identity"),
        data=data,
        backend="cmdstanr",
        seed=42,
        prior=priorBB
        )

  saveRDS(mod_distBB, file = here("R_output", "model_distanceBetaB.RDS"))
  }
```

```{r}
pp_check(mod_distBB,"stat_2d")

if (file.exists(here("R_output", "KfoldCV_distBetaB.RDS"))){
  KCV_distBB <- readRDS(here("R_output","KfoldCV_distBetaB.RDS"))
  }else{
    set.seed(42)
    KCV_distBB<-kfold(mod_distBB,K=10)
    saveRDS(KCV_distBB,file=here("R_output","KfoldCV_distBetaB.RDS"))
    }
```


```{r}
loo_compare(KCV_distB,KCV_distBB)
```

```{r}
buffer_widths <- c(100,300,600,900,1200,1500,1800)
```

we make all the binomial models for the urbanisation variables

```{r}
for(i in 1: length(buffer_widths)){

if (file.exists(here("R_output", paste0("model_",buffer_widths[i],"B.RDS")))){
    temp_mod <- readRDS(here("R_output", paste0("model_",buffer_widths[i],"B.RDS")))
  } else {  
  
mu_formula <- as.formula(
  paste0("N_blackleg | trials(N_total) ~ scaled_IMD_",buffer_widths[i],
                         " + (1 | SITE) + (scaled_IMD_",buffer_widths[i]," | CAMP)")
)


temp_mod <- brm(
  bf(mu_formula),
        family=binomial,
        data=data,
        backend="cmdstanr",
        seed=42,
        prior=priorB,
  control=list(adapt_delta=0.9)
        )
 
saveRDS(temp_mod, file = here("R_output", paste0("model_",buffer_widths[i],"B.RDS")))
  }
  

assign(paste0("mod_",buffer_widths[i],"B"), temp_mod)


}
```

then we do the same with all the beta_binomial

```{r}
for(i in 1: length(buffer_widths)){

if (file.exists(here("R_output", paste0("model_",buffer_widths[i],"BetaB.RDS")))){
    temp_mod <- readRDS(here("R_output", paste0("model_",buffer_widths[i],"BetaB.RDS")))
  } else {  
  
mu_formula <- as.formula(
  paste0("N_blackleg | trials(N_total) ~ scaled_IMD_",buffer_widths[i],
                         " + (1 | SITE) + (scaled_IMD_",buffer_widths[i]," | CAMP)")
)


temp_mod <- brm(
  bf(mu_formula,
     nlf(phi~1/invphi),
     invphi~1
     ),
        family=beta_binomial(link_phi="identity"),
        data=data,
        backend="cmdstanr",
        seed=42,
        prior=priorBB,
  control=list(adapt_delta=0.9)
        )
 
saveRDS(temp_mod, file = here("R_output", paste0("model_",buffer_widths[i],"BetaB.RDS")))
  }
  

assign(paste0("mod_",buffer_widths[i],"BB"), temp_mod)


}
```

let's do cross val for all these models

```{r}
for(i in 1: length(buffer_widths)){

if (file.exists(here("R_output", paste0("KfoldCV_",buffer_widths[i],"B.RDS")))){
    temp_KCV <- readRDS(here("R_output", paste0("KfoldCV_",buffer_widths[i],"B.RDS")))
  } else {  

    temp_mod <- readRDS(here("R_output", paste0("model_",buffer_widths[i],"B.RDS")))  
set.seed(42)
    temp_KCV<-kfold(temp_mod,K=10,
                    model_names = paste0("mod_",buffer_widths[i],"B"))
    
 
saveRDS(temp_KCV, file = here("R_output", paste0("KfoldCV_",buffer_widths[i],"B.RDS")))
  }
  

assign(paste0("KCV_",buffer_widths[i],"B"), temp_KCV)


}
```

```{r}
for(i in 1: length(buffer_widths)){

if (file.exists(here("R_output", paste0("KfoldCV_",buffer_widths[i],"BetaB.RDS")))){
    temp_KCV <- readRDS(here("R_output", paste0("KfoldCV_",buffer_widths[i],"BetaB.RDS")))
  } else {  

    temp_mod <- readRDS(here("R_output", paste0("model_",buffer_widths[i],"BetaB.RDS")))  
set.seed(42)
    temp_KCV<-kfold(temp_mod,K=10,
                    model_names = paste0("mod_",buffer_widths[i],"BB"))
    
 
saveRDS(temp_KCV, file = here("R_output", paste0("KfoldCV_",buffer_widths[i],"BetaB.RDS")))
  }
  

assign(paste0("KCV_",buffer_widths[i],"BB"), temp_KCV)


}
```

```{r}
### to explore
performance::variance_decomposition(mod_100BB)
performance::r2_nakagawa(mod_100BB)



varcomps_dist <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_distBB, re_formula = NA)),
  VI = VarCorr(mod_distBB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "dist",
    dist=2200,
    type="dist",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_dist$V_explained)

varcomps_100 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_100BB, re_formula = NA)),
  VI = VarCorr(mod_100BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "100",
    dist=100,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_100$V_explained)

varcomps_300 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_300BB, re_formula = NA)),
  VI = VarCorr(mod_300BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "300",
    dist=300,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_300$V_explained)

varcomps_600 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_600BB, re_formula = NA)),
  VI = VarCorr(mod_600BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "600",
    dist=600,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_600$V_explained)

varcomps_900 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_900BB, re_formula = NA)),
  VI = VarCorr(mod_900BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "900",
    dist=900,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_900$V_explained)

varcomps_1200 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_1200BB, re_formula = NA)),
  VI = VarCorr(mod_1200BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "1200",
    dist=1200,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_1200$V_explained)

varcomps_1500 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_1500BB, re_formula = NA)),
  VI = VarCorr(mod_1500BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "1500",
    dist=1500,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_1500$V_explained)

varcomps_1800 <- tibble(
  VF = matrixStats::rowVars(posterior_linpred(mod_1800BB, re_formula = NA)),
  VI = VarCorr(mod_1800BB, summary = FALSE)$SITE$sd[, "Intercept"]^2
) %>%
  mutate(
    model = "1800",
    dist=1800,
    type="IMD",
    V_SITE = VF + VI
  )%>%
  mutate(V_explained = VF/V_SITE)

mean_hdi(varcomps_1800$V_explained)


rbind(varcomps_dist,
      varcomps_100,
      varcomps_300,
      varcomps_600,
      varcomps_900,
      varcomps_1200,
      varcomps_1500,
      varcomps_1800) |> 
  ggplot()+
  stat_eye(aes(dist,V_explained*100),.width=c(0.001,0.95)) +
  geom_vline(xintercept=2000,lty=2)+
  scale_x_continuous("urbanisation metric",breaks=c(100,300,600,900,1200,1500,1800,2200),
                     labels=c("IMD \n 100m","IMD \n 300m","IMD \n 600m",
                              "IMD \n 900m","IMD \n 1200m","IMD \n 1500m","IMD \n 1800m",
                              "distance to centre"))+
  scale_y_continuous("% of (logit scale) among-site variance explained")+
  theme_bw()
```

```{r}
## looking at seasonal variation of urbanisation effect relative to grand mean
as_draws_df(mod_100BB) |> select(b_scaled_IMD_100,sd_CAMP__scaled_IMD_100) |> mutate(CV=sd_CAMP__scaled_IMD_100/abs(b_scaled_IMD_100)) |> mean_hdi(CV)
```



```{r}

tibble(
  `distance to centre`=-as_draws_df(mod_distBB)$b_scaled_dist, #inverted because values go in opposite direction
  `IMD (100m)`=as_draws_df(mod_100BB)$b_scaled_IMD_100,
  `IMD (300m)`=as_draws_df(mod_300BB)$b_scaled_IMD_300,
  `IMD (600m)`=as_draws_df(mod_600BB)$b_scaled_IMD_600,
  `IMD (900m)`=as_draws_df(mod_900BB)$b_scaled_IMD_900,
  `IMD (1200m)`=as_draws_df(mod_1200BB)$b_scaled_IMD_1200,
  `IMD (1500m)`=as_draws_df(mod_1500BB)$b_scaled_IMD_1500,
  `IMD (1800m)`=as_draws_df(mod_1800BB)$b_scaled_IMD_1800
) |> 
  pivot_longer(cols=everything(),values_to="effect",names_to="model") |> 
  ggplot()+
  stat_eye(aes(model,effect))+
  geom_hline(yintercept = 0,lty=2)+
  scale_x_discrete("urbanisation model")+
  scale_y_continuous("effect size of increasing urbanisation")

```

# Figures

```{r}
data_pooled = data |>  group_by(SITE) |> summarise(N_total=sum(N_total),N_blackleg=sum(N_blackleg)) |> left_join(sites)
```

```{r}
newdata=data |> 
  select(meanIMD_100m,scaled_IMD_100)  |>  
  distinct()  |>  
  mutate(N_total=1) |> 
  add_epred_draws(mod_100BB,re_formula=NA)

data |>  
  ggplot()+
  geom_point(aes(meanIMD_100m,N_blackleg/N_total,
                 size=N_total),col="grey")+
  geom_point(data=data_pooled,aes(meanIMD_100m,N_blackleg/N_total,
                 size=N_total))+
  stat_lineribbon(data=newdata,aes(x=meanIMD_100m,y=.epred),
                  .width=c(0.001,0.95),fill="grey",alpha=0.5)+
  scale_x_continuous("mean Imperviousness Density (in 100 m radius around site)")+
  scale_y_continuous("proportion of black-legged individuals")+
  scale_size("Beetles captured")+
  theme_bw()



newdata=data |> 
  select(meanIMD_100m,scaled_IMD_100,CAMP)  |>  
  mutate(SITE="RB00") |> 
  distinct()  |>  
  mutate(N_total=1) |> 
  add_epred_draws(mod_100BB,re_formula=~(scaled_IMD_100|CAMP))

data |>  
  ggplot()+
  geom_point(aes(meanIMD_100m,N_blackleg/N_total,
                 size=N_total),col="grey")+
  stat_lineribbon(data=newdata,aes(x=meanIMD_100m,y=.epred),
                  .width=c(0.001,0.95),fill="grey",alpha=0.5)+
  scale_x_continuous("mean Imperviousness Density (100 m radius around site)")+
  scale_y_continuous("proportion of black-legged individuals")+
  theme_bw()+facet_wrap(~CAMP)
```

as a supplementary figure, to bolster the case that no evidence of within year seasonal variation, so doy and year not included in study (between campaign within site var is assumed to be random)

```{r}
ggplot(data)+
    geom_point(aes(doy,N_blackleg/N_total,
                   size=N_total,col=YEAR))+facet_wrap(~SITE)
```

```{r}
sampling_effort <-dates |> 
  mutate(YEAR=factor(YEAR),
         doy_start=yday(dmy(DATE_start)),
         doy_end=yday(dmy(DATE_end))) |> 
           group_by(YEAR) |> 
  summarise(sampling_effort=max(doy_end)-min(doy_start))

data_pooled2 = data |>  group_by(SITE, YEAR) |> 
  summarise(N_total=sum(N_total),
            N_blackleg=sum(N_blackleg)) |> 
  left_join(sites) |> 
  left_join(sampling_effort)
```

no evidence that popsize, measure
```{r}
prior=c(
          set_prior("normal(0,1)",class="b"),   ###need better prior for intercept, or rate
          set_prior("normal(0,1)",nlpar="invshape",lb=0)
        )

mod_100 <- brm(
  bf(N_total|rate(sampling_effort)~0+YEAR+scaled_IMD_100:YEAR,
     nlf(shape~1/invshape),
     invshape~1),
        family=negbinomial(link_shape="identity"),
        data=data_pooled2,
        backend="cmdstanr",prior=prior,
        seed=42
        )
```

```{r}
data$Ndays=data$doy_end-data$doy_start

priorP=c(
          set_prior("normal(0,1)",class="b"),
          set_prior("normal(0,1)",class="Intercept"),
          set_prior("normal(0,1)",class="sd")
        )

mod=brm(N_total|rate(Ndays)~scaled_IMD_1800+(1|SITE)+(scaled_IMD_1800|CAMP),
        data=data,family=negbinomial,
        backend="cmdstanr",
        seed=42,
        prior=priorP,
  control=list(adapt_delta=0.9)
)
```

